Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks

### 2016-02-24 gm_newphyt/manuscript_gmes.Rmd
updates to gmes manuscript from SVC

\
The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients. As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co]. Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic]. Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization]. Higher rates of *A~n~* and stomatal conductance (g~s~) can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal]. Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].

@@

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).

Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`). Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).

Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively). We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~. Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.

@@

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```  
\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c). 
\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3**. The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals for the mean.
\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/physiology_boxplots.R")
```  
\
**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
source("master_scripts/wateruse.R")
```  
\
**Figure 5**. (a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
\
```{r gmshade, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gmes_shadeleaves_figure.R")
```  
\
**Figure 6**. The change in measured g~m for individual shade leaves following an increase in PPFD to match the light environment of the full sun leaf in the same tree. Each measurement of g~m~ on the same leaf, for both low and high environments, was recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette. The dashed line is the 1:1 relationship.

>>>>

\
The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients. As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does]. Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic]. Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization]. Higher rates of *A~n~* and stomatal conductance (g~s~) can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal]. Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].

@@

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001). Following the increase in PPFD from low to high light on the same leaf, g~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`). These measured values of g~m~ for shade leaves at high light were also equivalent or greater than those of sun leaves (Table `r wtc3tab2`). Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively). We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~. Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.

@@

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```  
\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c). 
\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3**. The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals for the mean.
\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/physiology_boxplots.R")
```  
\
**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
source("master_scripts/wateruse.R")
```  
\
**Figure 5**. (a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
\
```{r gmshade, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gmes_shadeleaves_figure.R")
```  
\
**Figure 6**. The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree. Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased. The dashed line is the 1:1 relationship.
